ABX usage

### 2019-12-11 scripts/27_maASlin2.Rmd
update

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Maaslin2) 
library(eeptools)
library(tidyverse)
```

@@

@@

# the three ILs, select the highest value if there are multiple 
pheno <- pheno %>% 
  # get patient info such as age and sex (using MRN)
  left_join(ctc.patients %>% 
              dplyr::select(MRN, sex, dob), by = 'MRN') %>% 
  # get the hightest ferritin value 
  left_join(ctc.ferritin %>% 
              dplyr::select(MRN, result_text) %>% 
              group_by(MRN) %>% 
              summarise(ferritin = max(result_text)), by = 'MRN') %>% 
  left_join(ctc.crp %>% 
              dplyr::select(MRN, result_text) %>% 
              mutate(result_text = if_else(str_detect(result_text, '^<'), '0', result_text),
                     result_text = as.numeric(result_text)) %>% 
              group_by(MRN) %>% 
              summarise(crp  = max(result_text, na.rm = T)), by = 'MRN') %>% 
  left_join(IL_table, by  ='MRN')

@@

```{r}
# now change the columns to different type: dbl or factor
pheno_final  <- pheno %>%  
  # change the ferritin to dbl
  mutate(ferritin = as.numeric(ferritin),
         Key__Patient_Disease_Status_at_CAR = as.character(Key__Patient_Disease_Status_at_CAR)) %>% 
  # change all the pheno vars to factors
  mutate_if(is.character, as.factor) %>% 
  mutate(MRN = as.character(MRN),
         sampleid = as.character(sampleid)) %>% 
  mutate(IL_6 = log(IL_6),
         IL_10 = log(IL_10)) %>% 
  column_to_rownames('sampleid')

colnames(pheno_final)
```

@@

fit_data <- Maaslin2(
    pw_fil, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = T,
    fixed_effects = c('Toxicity', 
                      'Cytokine_Release_Syndrome',
                      'ICANS',
                      'CR',
                      'ORR', 
                      'Disease_Class',
                      'Key__Patient_Disease_Status_at_CAR',
                      'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
                      'ferritin',
                      'crp',
                      'IL_1_beta',
                      'IL_10',
                      'IL_6'),
    random_effects = c('sex',
                       'age'),
    standardize = FALSE)

@@

```{r}
trans <- log(pw[,'PWY-7456'] + 1)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Maaslin2) 
library(eeptools)
library(tidyverse)
library(arrangements)
```

@@

```{r}
# add the lines of therapy var
lines <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
                    sheet = 8)  %>% 
  rename_all(funs(str_replace_all(., ' ','_'))) %>% 
  dplyr::select(MRN, Prior_lines_of_therapy)
```

@@

# the three ILs, select the highest value if there are multiple 
pheno <- pheno %>% 
  # get patient info such as age and sex (using MRN)
  left_join(ctc.patients %>% 
              dplyr::select(MRN, sex, dob), by = 'MRN') %>% 
  # get the hightest ferritin value 
  left_join(ctc.ferritin %>% 
              dplyr::select(MRN, result_text) %>% 
              group_by(MRN) %>% 
              summarise(ferritin = max(result_text)), by = 'MRN') %>% 
  left_join(ctc.crp %>% 
              dplyr::select(MRN, result_text) %>% 
              mutate(result_text = if_else(str_detect(result_text, '^<'), '0', result_text),
                     result_text = as.numeric(result_text)) %>% 
              group_by(MRN) %>% 
              summarise(crp  = max(result_text, na.rm = T)), by = 'MRN') %>% 
  left_join(IL_table, by  ='MRN') %>% 
  left_join(lines, by = 'MRN')

@@

```{r pheno_final}
# now change the columns to different type: dbl or factor
pheno_final  <- pheno %>%  
  # change the ferritin to dbl
  mutate(ferritin = as.numeric(ferritin),
         Key__Patient_Disease_Status_at_CAR = as.character(Key__Patient_Disease_Status_at_CAR)) %>% 
  # change all the pheno vars to factors
  mutate_if(is.character, as.factor) %>% 
  mutate(MRN = as.character(MRN),
         sampleid = as.character(sampleid)) %>% 
  #mutate(IL_6 = log(IL_6),
  #       IL_10 = log(IL_10)) %>% 
  column_to_rownames('sampleid')

colnames(pheno_final)

pheno_final %>% 
  summary
```

```{r}
# check the correlation between the pheno vars that are factors
all_var <- pheno_final %>% 
  select_if(is.factor) %>% 
  names

combi <- combinations(all_var, 2) %>% 
  as.data.frame(stringsAsFactors = F)

phe_cor <- apply(combi, 1, function(Row) {
  phe_fil = pheno_final %>% 
    dplyr::select(Row[['V1']], Row[['V2']])

  round(cor.test(pheno_final[,Row[['V1']]],  pheno_final[,Row[['V2']]],method = 'spearman', 
                           exact = F)$estimate, 2)
  #chisq.test(phe_fil)$p.value
}) %>% 
  as.data.frame() %>% 
  bind_cols(combi) %>% 
  mutate(phe_var = str_glue('{V1}__{V2}')) %>% 
  dplyr::select(-V1, -V2)

@@

fit_data <- Maaslin2(
    pw, pheno_final, '../data/maASlin2/pw', 
    normalization  = 'NONE', 
    transform = "LOG",  
    cores = 4, 
    plot_heatmap = T,
    min_prevalence = 0.5,
    fixed_effects = c('Toxicity', 
                      'CR',
                      'Disease_Class',
                      'Key__Patient_Disease_Status_at_CAR',
                      'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
                      'ferritin',
                      'crp',
                      'IL_1_beta',
                      'IL_10',
                      'IL_6',
                      'Prior_lines_of_therapy'),
    random_effects = c('sex',
                       'age'),
    standardize = FALSE,
    max_significance = 0.1)

@@

### 2019-12-11 scripts/29_abx_usage.Rmd
update

---
title: "Humann2 bar plot for gene family"
output: html_document
---

@@

######################################## Bash ########################################

```{bash}
# needs to download the utility mapping db first
humann2_databases --download utility_mapping full /Users/daia1/melodyP/data/humann2_res
```

```{bash}
# regroup to KEGG Orthogroups (KOs)
humann2_regroup_table \
            --input ../data/humann2_res/humann2_genefamilies.tsv  \
            --groups uniref90_ko  \
            --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
```

```{bash}
# renorm to relab
humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u relab --output ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv
```

```{bash}
# split into unstratified and ..
humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv  --output ../data/humann2_res/genefamily_split_KO
```

######################################## R ########################################

## What features(gene orthologs) do you wanna see? Do lefse first.

```{r}
# load the lefse res of tox
# the input : lefse res files
fns <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$', full.names = T)
fns_name <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$')

# join all of the tables feature together
feature <- fns %>% 
  set_names(fns_name) %>% 
  # when meeting parsing failures, you need to specify all column types.
  map(~ read_tsv(., col_names = c('genefam','xx','direction','score','pval'),  col_types = 'ccccc') %>% 
        filter(!is.na(score))) %>% 
  bind_rows(.id = 'group') %>% 
  mutate(group = str_replace(group, 'genefam_abundace_pheno__','')) %>% 
  mutate(group = str_replace(group, '.res$','')) %>% 
  select(-xx, -pval)

# change the "N" direction to be minus score
# feature <- bind_rows(
#   feature %>% 
#     split(.$direction) %>% 
#     pluck('N') %>% 
#     mutate(score = -score),
#   feature %>% 
#     split(.$direction) %>% 
#     pluck('Y') 
# ) %>% 
#   arrange(group, pathway, score)

feature %>% 
  write_csv('../data/lefse_genefam/with_unmapped_sig_feature.csv')

>>>>

---
title: "ABX usage"
output: html_document
---

@@

```{r}
